Inferring the gravitational potential of the Milky Way with a few precisely measured stars

### 2014-03-16 text/paper2/ms.tex
one sim

We leave the potential parameters --- $q_1$, $q_z$, $\phi$, and $\vhalo$ --- free to vary, along with the Lagrange point offset, $\Loffset$, and use Markov Chain Monte Carlo (MCMC) to sample from the posterior probability distribution over these five parameters. We use an affine-invariant ensemble sampler \citep{goodman10} implemented in the Python programming language \citep{foremanmackey13} with an ensemble of 64 walkers initialized by sampling from the priors summarized in Table~\ref{tbl:params}. We run the walkers in parallel (using MPI todo) for 5000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 10000 steps. We compute the autocorrelation time for each parameter [todo: reference] and thin the chains by taking every XXX$\times \mathrm{median}(t_{\rm acor})$ sample to ensure they are close to independent. Figure~\ref{fig:} shows the walker positions over the 10000 post-burn-in steps for each parameter for the $2.5\times10^8~M_\odot$ simulation; the autocorrelation time for each parameter is displayed on each corresponding panel. Figure~\ref{fig:}c shows the posterior probability for these five parameters for the same simulation.

\subsection{\spitzer + \gaia}
For this and subsequent experiments (unless specified), we use only the $2.5\times10^8~M_\odot$ simulation. We take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties. We take these stars to be RR Lyrae variables which are known to be excellent distance indicators via the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}. $\sim$100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%. These stars will also be included in the \gaia\, proper motion catalog: at a distance of $\sim$50~kpc, a typical RR Lyrae will have a tangential velocity error around $\sim$20~km/s.

>>>>

We leave the potential parameters --- $q_1$, $q_z$, $\phi$, and $\vhalo$ --- free to vary, along with the Lagrange point offset, $\Loffset$, and use Markov Chain Monte Carlo (MCMC) to sample from the posterior probability distribution over these five parameters. We use an affine-invariant ensemble sampler \citep{goodman10} implemented in the Python programming language \citep{foremanmackey13} with an ensemble of 64 walkers initialized by sampling from the priors summarized in Table~\ref{tbl:params}. We run the walkers in parallel (using MPI todo) for 5000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 10000 steps. We compute the autocorrelation time for each parameter [todo: reference] and thin the chains by taking every XXX$\times \mathrm{median}(t_{\rm acor})$ sample to ensure they are close to independent. Figure~\ref{fig:} shows the walker positions over the 10000 post-burn-in steps for each parameter for the $2.5\times10^8~M_\odot$ simulation; the autocorrelation time for each parameter is displayed on each corresponding panel. Figure~\ref{fig:} shows projections of the posterior probability distribution for the parameters.

\subsection{\spitzer + \gaia}
We take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties. We take these stars to be RR Lyrae variables which are known to be excellent distance indicators via the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}. $\sim$100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%. These stars will also be included in the \gaia\, proper motion catalog: at a distance of $\sim$50~kpc, a typical RR Lyrae will have a tangential velocity error around $\sim$20~km/s.
